---
title: "ingmangabriel_esm244_assignment1_task1"
format: 
  html:
    code-fold: true
    toc: true
    number-sections: true
editor: visual
execute:
  echo: true
  message: false
  warning: false
---

```{r setup}

library(tidyverse)
library(here)
library(tsibble)
library(feasts)
library(lubridate)
library(fable)
library(patchwork)

#install.packages('patchwork')

```

# Task 1

1.  A graph of total **mountain yellow-legged frog (Rana muscosa) count** each year across all water bodies, by life stage **excluding the ‘EggMass’ level**. 

2.  A column graph containing total counts (over all years of the study) of **combined adult and subadult endangered mountain yellow-legged frogs** (Rana muscosa) observed in the 5 lakes with the greatest total observed counts. 

**Combine your two graphs into a single compound figure using patchwork or cowplot.** 

-   Add a finalized figure caption that appears below the compound figure in your rendered html.  The caption should include a brief overall explanation of the multi-panel figure, and then for each part (A and B) a brief explanation of that panel.

-   **Note:** You might realize once you combine your graphs into a single compound figure, you need to update the formatting (e.g. move legends or instead directly label, etc.). You should customize as necessary to make it a professional final output.

### Task 1: Graph 1

```{r datawrangle}

amphibians_raw <- read.csv(here("data", "sierra_amphibians.csv"))
```

```{r filtergraph1}

amphibians_filtered <- amphibians_raw %>% 
  mutate(date = lubridate::mdy(survey_date)) %>% 
  filter(amphibian_species == 'RAMU') %>% 
  mutate(recorded_year = year(date)) %>% 
  filter(amphibian_life_stage == 'Tadpole' | amphibian_life_stage == 'Adult' | amphibian_life_stage == 'SubAdult') %>% 
  group_by(recorded_year, amphibian_life_stage) %>% 
  count()
```

```{r graph1}

frog_graph1 <- ggplot(data = amphibians_filtered, mapping = aes(x = recorded_year, y = n, group = amphibian_life_stage)) + 
  geom_line(aes(color = amphibian_life_stage)) +
  labs(title = 'Total Yellow-Legged Frog (Rana Muscosa) Count By Year', x = 'Year', y = 'Amphibian Count') +
  theme_light()
  
  
```

### Task 1: Graph 2

```{r filterdata2}

sierra_top_5 <- amphibians_raw %>% 
  filter(amphibian_species == 'RAMU') %>% 
  filter(amphibian_life_stage == 'Adult' | amphibian_life_stage == 'SubAdult') %>% 
  group_by(lake_id) %>% 
  summarise(n = n()) %>% 
  mutate(lake_string = paste('Lake', lake_id)) %>% #I won't pretend that the following code for this chunk was my idea. This was devised by a friend of mine who works at Fish and Wildlife on a Discord call. They are way better at R than me. This turns the lake_id column into a string that can be more easily called by other functions.
  slice_max(n, n = 5) %>% #selecting the lakes in the top 5
  mutate(lake_string = fct_reorder(lake_string, n, .fun = 'max'))

```

```{r graph2}

frog_graph2 <- ggplot(data = sierra_top_5, aes(x = lake_string, y = n)) +
  geom_col(
    colour = 'cornflowerblue'
  ) +
  labs(
    title = 'Top Five Lakes for Adult and SubAdult Rana Muscosa',
    x = 'Amphibian Count',
    y = 'Lake',
    cex.main = .25
  ) +
  theme_light()

frog_graph2

```
